### 2015-09-02 HarryPlotter/doc/harry_plotter.tex
new Artus/HarryPlotter illustrations

\begin{tikzpicture}[scale=1.0]
\node at (0cm,0cm) [draw, rectangle, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm] (module1) {Input Module 1};
\node [rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module1] (module2) {$\cdots$};
\node [draw, rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module2] (module3) {Analysis Module 1};
\node [draw, rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module3] (module4) {Analysis Module 2};
\node [draw, rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module4] (module5) {Analysis Module 3};
\node [rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module5] (module6) {$\cdots$};
\node [draw, rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module6] (module7) {Plot Module 1};
\node [rectangle, fill=white, minimum width=3.5cm, minimum height=5ex, node distance=5ex+0.3cm, below of=module7] (module8) {$\cdots$};

\node [draw, rectangle, rounded corners, KIT_rot_1, fill=KIT_rot_5, minimum width=3.5cm, minimum height=7ex, node distance=7ex+0.5cm, above of=module1] (input) {Input};
\node [draw, rectangle, rounded corners, KIT_blau_1, fill=KIT_blau_5, minimum width=3.5cm, minimum height=7ex, node distance=7ex+0.5cm, below of=module8] (plot) {Plot};

\draw [->, thick] (input) -- (module1);
\draw [->, thick] (module1) -- (module2);
\draw [->, thick] (module2) -- (module3);
\draw [->, thick] (module3) -- (module4);
\draw [->, thick] (module4) -- (module5);
\draw [->, thick] (module5) -- (module6);
\draw [->, thick] (module6) -- (module7);
\draw [->, thick] (module7) -- (module8);
\draw [->, thick] (module8) -- (plot);

\draw [fill=white] ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-5ex-1.5cm,2.5ex) $) rectangle node{\rotatebox{90}{Argument Parser}} ($ (module8) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,-2.5ex) $);

\draw [->, thick] ($ (module1) + (-1.75cm,1ex) $) -- ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,1ex) $);
\draw [->, thick] ($ (module3) + (-1.75cm,1ex) $) -- ($ (module3) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,1ex) $);
\draw [->, thick] ($ (module4) + (-1.75cm,1ex) $) -- ($ (module4) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,1ex) $);
\draw [->, thick] ($ (module5) + (-1.75cm,1ex) $) -- ($ (module5) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,1ex) $);
\draw [->, thick] ($ (module7) + (-1.75cm,1ex) $) -- ($ (module7) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm,1ex) $);

\draw [<-, thick, KIT_gruen_1] ($ (module1) + (-1.75cm,-1ex) $) -- ($ (module1) + (-1.75cm-0.5cm,-1ex) $);
\draw [<-, thick, KIT_gruen_1] ($ (module3) + (-1.75cm,-1ex) $) -- ($ (module3) + (-1.75cm-0.5cm,-1ex) $);
\draw [<-, thick, KIT_gruen_1] ($ (module4) + (-1.75cm,-1ex) $) -- ($ (module4) + (-1.75cm-0.5cm,-1ex) $);
\draw [<-, thick, KIT_gruen_1] ($ (module5) + (-1.75cm,-1ex) $) -- ($ (module5) + (-1.75cm-0.5cm,-1ex) $);
\draw [<-, thick, KIT_gruen_1] ($ (module7) + (-1.75cm,-1ex) $) -- ($ (module7) + (-1.75cm-0.5cm,-1ex) $);

\node at ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-0.5cm,1ex-1.75ex) $) {\textit{define}};
\node at ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-0.5cm,1ex-1.75ex-2.25ex) $) {\textit{arguments}};

\draw [rounded corners, KIT_gruen_1, fill=white] ($ (module1) + (-1.75cm-0.5cm-5ex,2.5ex) $) rectangle node{\rotatebox{90}{Settings}} ($ (module8) + (-1.75cm-0.5cm,-2.5ex) $);

\draw [<-, thick, KIT_gruen_1] ($ (module1) + (-1.75cm-0.5cm-2.5ex,2.5ex) $) -- ($ (module1) + (-1.75cm-0.5cm-2.5ex,2.5ex+0.3cm) $) -- ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-2.5ex,2.5ex+0.3cm) $) -- ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-2.5ex,2.5ex) $);
\draw [<-, thick, KIT_gruen_1] ($ (module8) + (-1.75cm-0.5cm-2.5ex,-2.5ex) $) -- ($ (module8) + (-1.75cm-0.5cm-2.5ex,-2.5ex-0.3cm) $) -- ($ (module8) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-2.5ex,-2.5ex-0.3cm) $) -- ($ (module8) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-2.5ex,-2.5ex) $);

\node [KIT_gruen_1] at ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-0.5cm,2.5ex+0.3cm+1.5ex) $) {\textit{1${}^\text{st}$ parsing}};
\node [KIT_gruen_1] at ($ (module8) + (-1.75cm-0.5cm-5ex-0.5cm-0.5cm,-2.5ex-0.3cm-1.5ex) $) {\textit{2${}^\text{nd}$ parsing}};

\draw [rounded corners, KIT_gruen_1, fill=KIT_gruen_5] ($ (module1) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-5ex-0.5cm-5ex,2.5ex) $) rectangle node{\rotatebox{90}{Program Arguments, JSON Configuration}} ($ (module8) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-5ex-0.5cm,-2.5ex) $);

\draw [->, thick, KIT_gruen_1] ($ (module4) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-5ex-0.5cm,-2.5ex-0.15cm) $) -- ($ (module4) + (-1.75cm-0.5cm-5ex-0.5cm-1.5cm-5ex,-2.5ex-0.15cm) $);

\draw [rounded corners, KIT_rot_1, fill=white] ($ (module1) + (1.75cm+0.5cm+5ex,2.5ex) $) rectangle node{\rotatebox{90}{Plot Data}} ($ (module8) + (1.75cm+0.5cm,-2.5ex) $);

\draw [<->, thick, KIT_rot_1] (module1) -- ($ (module1) + (1.75cm+0.5cm,0cm) $);
\draw [<->, thick, KIT_rot_1] (module3) -- ($ (module3) + (1.75cm+0.5cm,0cm) $);
\draw [<->, thick, KIT_rot_1] (module4) -- ($ (module4) + (1.75cm+0.5cm,0cm) $);
\draw [<->, thick, KIT_rot_1] (module5) -- ($ (module5) + (1.75cm+0.5cm,0cm) $);
\draw [<->, thick, KIT_rot_1] (module7) -- ($ (module7) + (1.75cm+0.5cm,0cm) $);
\end{tikzpicture}