### 2015-09-09 Core/doc/latex/include/bib.tex
paper: added bibliography

 \begin{thebibliography}{99}

    \bibitem{joram_phd}Joram Berger,
      \textit{Search for the Higgs Boson Produced via Vector-Boson
                        Fusion in the Decay Channel H $\to \tau \tau$},
      CERN-THESIS-2014-092.

\end{thebibliography}