### 2015-10-28 Core/doc/latex/include/bib.tex
Paper: Reorder citations

 \begin{thebibliography}{99}
    \bibitem{wlcg} \url{http://wlcg-public.web.cern.ch}.
    \bibitem{joram_phd}Joram Berger,
      \textit{Search for the Higgs Boson Produced via Vector-Boson
                        Fusion in the Decay Channel H $\to \tau \tau$},
      \href{https://cds.cern.ch/record/1747055}{\texttt{CERN-THESIS-2014-092}}.
    \bibitem{cmssw} \url{https://github.com/cms-sw/cmssw}
    \bibitem{boost} \url{http://www.boost.org}
    \bibitem{root}
      Rene Brun and Fons Rademakers, 
      \emph{ROOT -- An Object Oriented Data Analysis Framework}, 
      Proceedings AIHENP'96 Workshop, Lausanne, Sep. 1996, Nucl. Inst. \& Meth. in Phys. Res. A \textbf{389} (1997) 81--86. See also \url{http://root.cern.ch}.
    \bibitem{cms}
      CMS Collaboration,
      \emph{The CMS Experiment at the CERN LHC},
      JINST \textbf{3} (2008) S08004, 
      \textsc{doi}: \href{http://dx.doi.org/10.1088/1748-0221/3/08/S08004}{\texttt{10.1088/1748-0221/3/08/S08004}},
    \bibitem{kappa} \url{https://github.com/KappaAnalysis/Kappa}
\end{thebibliography}