### 2015-11-02 Core/doc/latex/figures/analysis_steps.tex

\begin{tikzpicture}[scale=1.0]
% horizontal axis
\draw[thick, ->] (0.7,0) -- (9.3,0) node[anchor=north] {};
% vertical axis
\draw[KIT_gruen_1,->] (0,0) -- (0,4) node[anchor=south] {File size /$\mathrm{MB}$};

\foreach \y/\ytext in {1/0,2/3, 3/6}
\draw[KIT_gruen_1,shift={(0,\y)}] (2pt,0pt)  node[left] {$10^\ytext$};

\filldraw[KIT_gruen_1] (0.3,3.11) circle(.05cm) node[black,anchor=south west]{Full event data};
\filldraw[KIT_gruen_1] (3.3,2.269) circle(.05cm) node[black, anchor=west]{Skim};
\filldraw[KIT_gruen_1] (6.6,1.03) circle(.05cm) node[black, anchor=south west]{N-Tuple};
\filldraw[KIT_gruen_1] (9.7,0) circle(.05cm) node[black, anchor=north]{Plot};

\draw[KIT_rot_1,<-] (10,0) -- (10,4) node[anchor=south] {Run time / h};
\foreach \y/\ytext in {1/-1,2/1, 3/3}
\draw[left,KIT_rot_1, shift={(0,\y)}, anchor=east] (11cm,0pt)  node[left] {\hfill$10^{\ytext}$};

\filldraw[KIT_rot_1] (1.65,0.66) circle(.05cm) node[black, anchor=west]{Skimming};
\filldraw[KIT_rot_1] (5,1.65) circle(.05cm) node[black, anchor=north]{Analysis};
\filldraw[KIT_rot_1] (8.3,3.08) circle(.05cm) node[black, anchor=south]{Plotting};
% file size
\draw[KIT_rot_1,->,thick] (0.5,0.3) -- (9.5,3.3) ;
\draw[KIT_gruen_1,->,thick] (.5,3.1) -- (9.5,0.2) ;

% labels
\draw (5,0) node[anchor=north] {Analysis Progress};
\end{tikzpicture}